---
title: "MA206 Code Movies"
subtitle: "A companion for the MA206 Tidyverse Tutorial"
author: "Code adapted by Conner Surrency and Gina Reynolds"
output: 
  rmdformats::readthedown:
    highlight: kate
---

```{r setup, include = F}
options(knitr.duplicate.label = "allow")
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, 
                      comment = "", cache = F, error = F)
library(flipbookr)
library(tidyverse)
library(gapminder)
```

---

#### "MA 206 Code Movies" is a companion guide to the MA206 tidyverse tutorial. Specifically, it dissects the Tutorial's and presents them *step-by-step*. 

#### For discussion of the example data and functions, refer back to the MA206 Tidyverse Tutorial guide.

###  
***
## 1. Using ***Tidyverse*** to Explore the Data
***

### Introduced in this example:

#### `read_csv()` read in *comma separated values* file 
#### `select()` choose columns
#### `filter()` keep or drop rows 
#### `mutate()` create new variable (column)
#### `summarise()` summarize data (sometimes used with `group_by()`)
####    

#### First, we use the *read_csv()* command to bring our data into RStudio, then look step-by-step look at what the *filter()*, *select()*, and *mutate()* commands can do in a pipeline:

```{r ex1, include=FALSE}
read_csv("wage_data.csv") ->
  wage_data

wage_data %>%
  filter(Occupation != "53: Never Worked" & 
           Earnings > 0) %>%  
  mutate(Earnings_per_member = Earnings / FamilySize) %>% 
  select(Earnings, Earnings_per_member, Sex, Age, Education) %>% 
  filter(Earnings< 150000) ->
employed_under_150K
```

```{r, echo=FALSE}
embed_flipbook(chunk_name = "ex1", 
                       use_share_again = T,
                       font_size = 90,
                       title_page = T,
                       title = "Ex. 1: Data Exploration",
                       subtitle = "Tailoring Your Data with tidyverse",
                       widths = c(1,1),
                       use_embed_xaringan = T)
```  

#### We can also use the *summarize()* command to create new variables for later use. Here, we use it to identify the mean earnings of our sample population:

```{r summarize, include=FALSE}
wage_data %>%
  filter(Occupation != "53: Never Worked" & 
           Earnings > 0) %>% 
  summarise(mean(Earnings)) %>% 
  as.numeric() ->
mean_earnings

#printing saved object
mean_earnings

wage_data %>%
  filter(Occupation != "53: Never Worked" & 
           Earnings > 0) %>% 
  group_by(Sex) %>% 
  summarise(mean = mean(Earnings),
            count = n()) ->
mean_earnings_by_sex

mean_earnings_by_sex
```

---

###      
***
## 2. Using ***ggplot2*** to Explore the Data
***

### Introduced in this example:

#### ggplot(), aes(), labs(), group_by(), facet_wrap(), geom_histogram(), geom_point(), geom_smooth(), geom_line(), geom_boxplot()
####      {.tabset .tabset-fade .tabset-pills}

##### Ex. 2.1: Histogram
```{r ex2.1, include = F}
employed_under_150K %>%
  ggplot() + 
  aes(x = Earnings) +
  geom_histogram(position = "dodge") +
  aes(fill = Sex) + 
  labs(x = "Earnings (USD)") +
  labs(y = "Count") + 
  labs(title = "Histogram of Earnings by Sex") +
  facet_wrap(facets = vars(Sex))
```

View code-movie [in a new tab](embedded_flipbooks/ex2.1_embed.html){target="_blank"}

```{r, echo = F}
embed_flipbook(chunk_name = "ex2.1", 
                       use_share_again = T,
                       font_size = 90,
                       title_page = T,
                       title = "Ex. 2.1: Building Your Plot",
                       subtitle = "Histogram",
                       widths = c(1,1),
                       use_embed_xaringan = T)
```  

##### Ex. 2.2: Scatterplot

```{r ex2.2, include = F}
employed_under_150K %>%
  ggplot() + 
  aes(x = Age) +
  aes(y = Earnings) +
  aes(color = Sex) +
  geom_point(alpha = 0.2) +
  labs(x = "Age (Years)") +
  labs(y = "Count") + 
  labs(title = "Age vs. Earnings by Sex") +
  facet_wrap(facets = vars(Sex)) +
  geom_smooth(color = "black")
```

```{r, echo = F}
embed_flipbook(chunk_name = "ex2.2", 
                       use_share_again = T,
                       font_size = 90,
                       title_page = T,
                       title = "Ex. 2.2: Building Your Plot",
                       subtitle = "Scatterplot",
                       widths = c(1,1),
                       use_embed_xaringan = T)
```  

##### Ex. 2.3: Line Plot

```{r ex2.3, include = F}
employed_under_150K %>% 
  group_by(Age) %>% 
  summarise(Median = median(Earnings)) %>% 
  ggplot() + 
  aes(x = Age) +
  aes(y = Median) +
  geom_line() +
  labs(x="Age (Years)") + 
  labs(y="Earnings (USD)") +
  labs(title = "Age vs. Median Earnings")

employed_under_150K %>% 
  group_by(Age, Sex) %>% 
  summarise(Median = median(Earnings)) %>% 
  ggplot() + 
  aes(x = Age) +
  aes(y = Median) +
  aes(color = Sex) +
  geom_line() +
  labs(x="Age (Years)") + 
  labs(y="Earnings (USD)") +
  labs(title = "Age vs. Median Earnings by Sex")
```

```{r, echo = FALSE}
embed_flipbook(chunk_name = "ex2.3", 
                       use_share_again = T,
                       font_size = 90,
                       title_page = T,
                       title = "Ex. 2.3: Building Your Plot",
                       subtitle = "Line Plot",
                       widths = c(1,1),
                       chunk_options = "fig.height = 3",
                       use_embed_xaringan = T)
``` 

---

###      
***
## 3. Using ***ggplot2*** functions for data visualization
***

### Introduced in this example:

#### labs() *[additional uses]*, scale_x_continuous(), scale_y_continuous(), limits(), breaks(), theme_X()
####    

```{r ex3, include = F}
employed_under_150K %>% 
  group_by(Age, 
           Sex) %>% 
  summarise(Median = median(Earnings)) %>% 
  ggplot(aes(x = Age, 
             y=Median)) +
  geom_line() +
  aes(color = Sex) +
  labs(x="Age (Years)") +
  labs(y="Earnings (USD)") +
  labs(color = "Sex of respondent") +
  labs(title = "Age vs. Median Earnings by Sex") +
  scale_x_continuous(limits=c(0,100), 
                     breaks=c(seq(from = 0,
                                  to = 100,
                                  by = 10))) +
  scale_y_continuous(limits=c(0,70000), 
                     breaks=c(10000, 30000, 50000)) +
  theme_bw() +
  theme_dark() +
  theme_light() +
  theme_gray() +
  theme_void() +
  theme_linedraw() +
  theme_minimal() +
  theme_classic()
```

```{r, echo = FALSE}
embed_flipbook(chunk_name = "ex3", 
                       use_share_again = T,
                       font_size = 90,
                       title_page = T,
                       title = "Ex. 3: More ggplot2 Commands",
                       subtitle = "Thematic and Display Adjustments",
                       widths = c(1,1),
                       use_embed_xaringan = T)
```     

---

###      
***
## 4. Additional ***ggplot2*** Geometric Objects
***

### Introduced in this example:

#### geom_col(), geom_bar(), geom_dotplot(), group_by(), factor(), c() 
####     {.tabset .tabset-fade .tabset-pills}

##### Ex. 4.1: Column Chart

```{r ex4.1, include = F}
employed_under_150K %>%
  group_by(Education, Sex) %>%
  summarize(mean_earnings = mean(Earnings)) %>%
  mutate(Education = factor(Education, 
                            levels = c("Less than HS", 
                                       "HS Graduate", 
                                       "Some College/Associates", 
                                       "Bachelors", 
                                       "Masters", 
                                       "PhD/MD/JD"))) %>% 
  ggplot() + 
  aes(x = mean_earnings, y = Education) + 
  aes(fill = Sex) + 
  geom_col(position = "dodge") + 
  labs(x = "Mean Earnings (USD)") + 
  labs(y = "Count") + 
  labs(title = "Column Chart of Mean Earnings by Education",
       subtitle = "Divided by Share of Each Sex") + 
  scale_x_continuous(labels = scales::comma)
```

```{r, echo = F}
embed_flipbook(chunk_name = "ex4.1", 
                       use_share_again = T,
                       font_size = 90,
                       title_page = T,
                       title = "Ex. 4.1: Column Chart",
                       widths = c(1,1),
                       use_embed_xaringan = T)
```


##### Ex. 4.2: Bar Plot
```{r ex4.2, include = F}
employed_under_150K %>%
  ggplot() + 
  aes(x = Sex) +
  aes(fill = Sex) +
# bar is a special case of geom col
# It counts the number of observations
# And this is the y ie height of the column
  geom_bar() + 
  labs(x = "Earnings (USD)") +
  labs(y = "Count") + 
  labs(title = "Bar Plot of Total Earnings by Sex")
```

```{r, echo = F}
embed_flipbook(chunk_name = "ex4.2", 
                       use_share_again = T,
                       font_size = 90,
                       title_page = T,
                       title = "Ex. 4.2: Bar Plot",
                       widths = c(1,1),
                       use_embed_xaringan = T)
``` 

##### Ex. 4.3: Dot Plot

```{r ex4.3, include = F}
library(gapminder)
gapminder %>%  
  filter(year == 2002) %>%  
  ggplot() +  
  aes(x = lifeExp) +  
  aes(fill = continent) +
  theme_minimal() +  
  geom_dotplot(alpha = .5)  +
  geom_rug(col = "darkgrey") +  
  labs(x = "Life Expectancy (Years)") +
  labs(y = "Count") + 
  labs(title = "National Life Expectancy by Region",
       subtitle = "Data from 2002") +
  facet_wrap(facets = vars(continent)) +
  theme(legend.position = "none")
```

```{r, echo=FALSE}
embed_flipbook(chunk_name = "ex4.3", 
                       use_share_again = T,
                       font_size = 90,
                       title_page = T,
                       title = "Ex. 4.3: Dot Plot",
                       widths = c(1,1),
                       use_embed_xaringan = T)
```

***

###    
### END

 

---

```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 80%}

@media print {
  .has-continuation {
    display: block;
  }
}
```
